\documentclass{irsarticle}

\usepackage{url}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}


\begin{document}
\title{\bf Recuperación de información geográfica}
\author{Jorge J. Morgado Vega\qquad Roberto García Rodríguez}
\date{}
\maketitle

\institution{Facultad de Matemática y computación, Universidad de la Habana}

\begin{abstract}
	Resumen del artículo
\end{abstract}

\keywords{recuperación de información, información geográfica}

\newpage
\tableofcontents
\newpage

\input{characteristics.tex}
\input{architecture.tex}
\input{evaluation.tex}
\input{applications.tex}
\input{pros_and_cons.tex}
\input{challenges.tex}


\begin{thebibliography}{}

\bibitem{chang2016}
Chang, K. T. (2016). Geographic information system. International Encyclopedia
of Geography: People, the Earth, Environment and Technology: People, the Earth,
Environment and Technology, 1-9.

\bibitem{teevan2011}
Teevan, J., Karlson, A., Amini, S., Brush, A. B., \& Krumm, J. (2011, August).
Understanding the importance of location, time, and people in mobile local
search behavior. In Proceedings of the 13th international conference on human
computer interaction with mobile devices and services (pp. 77-80).

\bibitem{leszczynski2012}
Leszczynski, A. (2012). Situating the geoweb in political economy.
Progress in human geography, 36(1), 72-89.

\bibitem{johnson2015}
Johnson, P. A., Corbett, J. M., Gore, C., Robinson, P., Allen, P., \& Sieber,
R. (2015). A web of expectations: Evolving relationships in community
participatory Geoweb projects. ACME: An International Journal for Critical
Geographies, 14(3), 827-848.

\bibitem{lymberopoulos2011}
Lymberopoulos, D., Zhao, P., Konig, C., Berberich, K., \& Liu, J.
(2011, October). Location-aware click prediction in mobile local search. In
Proceedings of the 20th ACM international conference on Information and
knowledge management (pp. 413-422).

\bibitem{turner2006}
Turner, A. (2006). Introduction to neogeography. " O'Reilly Media, Inc.".

\bibitem{connors2012}
Connors, J. P., Lei, S., \& Kelly, M. (2012). Citizen science in the age of neogeography: Utilizing volunteered geographic information for environmental monitoring. Annals of the Association of American Geographers, 102(6), 1267-1289.

\bibitem{harris2012}
Harris, T. M., \& Lafone, H. F. (2012). Toward an informal spatial data infrastructure: voluntary geographic information, neogeography, and the role of citizen sensors. SDI, Communities and Social Media, 8, 8-12.

\bibitem{artz2009}
ARTZ, M. (2009, September 14). Top Five Benefits of GIS. GIS AND SCIENCE.
\url{https://gisandscience.com/2009/09/14/top-five-benefits-of-gis/}

\bibitem{scholten1991}
Scholten, H. J., \& De Lepper, M. J. (1991). The benefits of the application of
geographical information systems in public and environmental health. World
health statistics quarterly. Rapport trimestriel de statistiques sanitaires
mondiales, 44(3), 160-170.

\bibitem{gey2007}
Gey, F., Larson, R., Sanderson, M., Bischoff, K., Mandl, T., Womser-Hacker, C.,
... \& Ferro, N. (2007). Challenges to evaluation of multilingual geographic
information retrieval in GeoCLEF. In quot; In Workshop on Evaluation of
Information Access (EVIA) May 15 (Tokyo Japan May 15 2007).

\bibitem{jones2008}
Jones, C. B., \& Purves, R. S. (2008). Geographical information retrieval.
International Journal of Geographical Information Science, 22(3), 219-228.

\bibitem{martins2005}
Martins, B., Silva, M. J., \& Chaves, M. S. (2005, November). Challenges and
resources for evaluating geographical IR. In Proceedings of the 2005 workshop
on Geographic Information Retrieval (pp. 65-69).

\bibitem{purves2011}
Purves, R., \& Jones, C. (2011). Geographic information retrieval. SIGSPATIAL
Special, 3(2), 2-4.

\bibitem{purves2018}
Purves, R. S., Clough, P., Jones, C. B., Hall, M. H., \& Murdock, V. (2018).
Geographic information retrieval: Progress and challenges in spatial search of
text. Foundations and Trends in Information Retrieval, 12(2-3), 164-318.

\bibitem{purves2014}
Purves, R. (2014). Geographic information retrieval: are we making progress. In
NCGIA specialist meeting on spatial search (pp. 1-6).

\bibitem{mandl2008}
Mandl, T., Gey, F., Nunzio, G. D., Ferro, N., Sanderson, M., Santos, D., \&
Womser-Hacker, C. (2008). An evaluation resource for Geographical Information
Retrieval. In quot; In Proceedings of the 6 th International Conference on
Language Resources and Evaluation (LREC 2008)(Marrakech 28-30 May 2008)
European Language Resources Association (ELRA). European Language Resources
Association (ELRA).

\bibitem{buscaldi2009}
Buscaldi, D. (2009, July). Toponym ambiguity in geographical information
retrieval. In Proceedings of the 32nd international acm sigir conference on
research and development in information retrieval (pp. 847-847).

\bibitem{cai2011}
Cai, G. (2011). Relevance ranking in geographical information retrieval.
SIGSPATIAL Special, 3(2), 33-36.

\bibitem{purves2004}
Purves, R., \& Jones, C. (2004, December). Workshop on geographic information
retrieval, SIGIR 2004. In ACM SIGIR Forum (Vol. 38, No. 2, pp. 53-56). New
York, NY, USA: ACM.

\bibitem{kumar2011}
Kumar, C. (2011, August). Relevance and ranking in geographic information
retrieval. In Fourth BCS-IRSG Symposium on Future Directions in Information
Access (FDIA 2011) 4 (pp. 2-7).

\bibitem{janowicz2011}
Janowicz, K., Raubal, M., \& Kuhn, W. (2011). The semantics of similarity in
geographic information retrieval. Journal of Spatial Information Science, (2),
29-57.

\bibitem{larson2004sep}
Larson, R. R., \& Frontiera, P. (2004, September). Spatial ranking methods for
geographic information retrieval (GIR) in digital libraries. In International
Conference on Theory and Practice of Digital Libraries (pp. 45-56). Springer,
Berlin, Heidelberg.

\bibitem{larson2004jul}
Larson, R. R., \& Frontiera, P. (2004, July). Ranking and representation for
geographic information retrieval. In Extended abstract in SIGIR 2004 Workshop
on Geographic Information Retrieval.

\bibitem{li1994}
Li, R. (1994). Data structures and application issues in 3-D geographic
information systems. Geomatica, 48(3), 209-224.

\bibitem{deAndrade2014}
de Andrade, F. G., de Souza Baptista, C., \& Davis, C. A. (2014). Improving
geographic information retrieval in spatial data infrastructures.
GeoInformatica, 18(4), 793-818.

\bibitem{abdelmoty2005}
Abdelmoty, A. I., Smart, P. D., Jones, C. B., Fu, G., \& Finch, D. (2005). A
critical evaluation of ontology languages for geographic information retrieval
on the Internet. Journal of Visual Languages \& Computing, 16(4), 331-358.

\bibitem{brown1999}
Brown, I. (1999). Developing a virtual reality user interface (VRUI) for
geographic information retrieval on the Internet. Transactions in GIS, 3(3),
207-220.

\bibitem{martins2007}
Martins, B., Borbinha, J., Pedrosa, G., Gil, J., \& Freire, N. (2007, November).
Geographically-aware information retrieval for collections of digitized
historical maps. In Proceedings of the 4th ACM Workshop on Geographical
information Retrieval (pp. 39-42).

\bibitem{laMancha}
Castilla-La Mancha. (2021, October 26). Wikipedia. Retrieved October 30, 2021,
from \url{https://es.wikipedia.org/wiki/Castilla-La_Mancha#Toponimia}.

\bibitem{purves2007} Purves, R. S., Clough, P., Jones, C. B., Arampatzis, A., Bucher, B.,
Finch, D., ... \& Yang, B. (2007). The design and implementation of SPIRIT: a
spatially aware search engine for information retrieval on the Internet.
International journal of geographical information science, 21(7), 717-745.

\bibitem{perea2007} Perea-Ortega, J. M., Cumbreras, M. Á. G., Vega, M. G., \&
Montejo-Ráez, A. (2007). GEOUJA System. University of Jaén at GeoCLEF 2007. In
CLEF (Working Notes).

\bibitem{bordogna2012} Bordogna, G., Ghisalberti, G., \& Psaila, G. (2012). Geographic
information retrieval: Modeling uncertainty of user's context. Fuzzy Sets and
Systems, 196, 105-124.

\end{thebibliography}

\end{document}
